---
title: "Kiribati BioMarkers Pipeline (Quarto Orchestrator)"
author: "Tinashe M. Tapera"
filters:
  - sorting-hat
  - ripper
extensions:
  sorting-hat:
    keep:
      - r
    verbose: true
  ripper:
    include-yaml: false
    output-dir: "../tests/testthat"
    output-name: "test-pipeline_orchestrator"
    script-links-position: "none"
format: gfm
---

This document is a Quarto Orchestrator pipeline that defines the workflow for 
processing the Kiribati BioMarkers dataset. The pipeline includes steps for data 
extraction, transformation, and loading (ETL) to prepare the dataset for analysis. 
Each step is defined as a separate `targets` target, and the pipeline ensures that 
tasks are executed in the correct order based on their dependencies. The pipeline 
is designed to be modular and reusable by utilizing `targets`, allowing for easy 
updates and maintenance as new data becomes available or as processing requirements change.

```{r}
library(here)
```

## Input Data

The first target is the input data. We staged input data to this project
using the `stagecoach` tool, which is a data staging and management tool that helps
organize and manage data for analysis. We track the stagecoach manifest
file as a dependency:

```{r}
#| format: file
#| packages: [here]

stage_coach_manifest <- here("stagecoach_manifest.yml")
```

Additionally, we can just track all of the files
in the input directory to track any changes:


```{r}
#| format: file
#| packages: [here]

raw_data <- list.files(
  here("data", "inputs"),
  full.names = TRUE,
  recursive = TRUE
)
```