---
title: "Data Validation (Alpha Version)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Validation (Alpha Version)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
#library(KiribatiBiomarkers)
library(dplyr)
library(skimr)
library(stringr)
library(readr)
library(lubridate)
library(targets)
library(here)
library(codebook)
library(labelled)

```

```{r, include=FALSE}
set.seed(1234)
```

This is a quick vignette to investigate and validate the alpha version of the dataset for distribution.

## What Data is Available

We've linked the data from GDrive to FASRC and symlinked it to the local `data/` directory using the `link_inputs()` function.

Now, we can take a look at the available files:

```{r}
tar_read(raw_data, store = here("_targets"))
```

There are several folders of data including:

1. Goldenetal_KI_biomarkers
2. Health data cleaning, exploration and merging to HIES
3. Kiribati_heavymetals
4. Kiribati-Mercury-Nails
5. ZamborainMasonetal_Seafood_KI

We will eventually go through each of these folders and validate the data within.

## 1. Goldenetal_KI_biomarkers

```{r}
golden_biomarkers_files <- tar_read(raw_data, store = here("_targets")) %>%
  str_subset(
    "Goldenetal_KI_biomarkers"
  ) %>%
  str_subset(
    "csv$"
  )
golden_biomarkers_files
```

```{r}
golden_biomarkers_data <- lapply(
  golden_biomarkers_files,
  readr::read_csv
)
names(golden_biomarkers_data) <- basename(golden_biomarkers_files)
```

We can now take a look at the dataframes loaded:

```{r}
lapply(
  golden_biomarkers_data,
  skimr::skim
)
```

> Addressing the email:

1) In the original data file, there was quite a bit of notes in column CE. Billy has now gone through those manually and entered them into the new data file that we will send on to collaborators. The Notes column is too much of a mess to send on to folks and we want something "clean-ish" and streamlined to be able to share with folks.

2) I'm seeing a data issue in column L. This should be someone's age in months, so essentially column I (date of interview) minus column J (date of birth) and then converted to months. Can we try to correct this? In order to properly de-identify people in the dataset we share, we should not be providing people's DOB which is why I only want age months provided.

3) I think we can remove the following columns when we share the data files. Can you confirm that none of these columns are essential for the code Jessica has written? I would like to remove the following

Columns: C-K (nothing that I & J will be replaced with age-months as mentioned above);  N-BK; BW-BY; ~~CA-CR~~ We will remove CA-CR BUT include CI-CS.

Column CE is column number 83:

```{r}
names(golden_biomarkers_data$kiribati_cleandata_mergedHIES.csv)[83] -> col_ce

golden_biomarkers_data$kiribati_cleandata_mergedHIES.csv %>%
  select(all_of(col_ce)) %>%
  summarise(n = n(), .by = all_of(col_ce))
```

These seem to tell us where the data was collected from, so can safely be removed.

To column L, which is 12:

```{r}
names(golden_biomarkers_data$kiribati_cleandata_mergedHIES.csv)[12] -> col_l
golden_biomarkers_data$kiribati_cleandata_mergedHIES.csv %>%
  select(all_of(col_l)) %>%
  summary()
```

This should instead be I minus J:

```{r}
names(golden_biomarkers_data$kiribati_cleandata_mergedHIES.csv) %>% str_subset("DOB_HIES$") -> col_dob
names(golden_biomarkers_data$kiribati_cleandata_mergedHIES.csv) %>% str_subset("date_of_interview") -> col_int_date

golden_biomarkers_data$kiribati_cleandata_mergedHIES.csv %>%
  mutate(
    dob = as.Date(.data[[col_dob]], format = "%d/%m/%Y"),
    interview_date = as.Date(.data[[col_int_date]], format = "%d/%m/%Y"),
    age_in_months = dob %--% interview_date / months(1)
  ) %>%
  select(all_of(col_dob), all_of(col_int_date), dob, interview_date, age_in_months) -> age_check
```

It looks like there are some people who are registered with DOBs that are after their interview date:

```{r}
age_check %>%
  filter(age_in_months <= 1)
```

How many people is this?

```{r}
nrow(age_check)
age_check %>%
  filter(age_in_months <= 1) %>%
  nrow()
```

4 of 1305 is not bad. I suggest removing them.

To remove columns C-K, N-BK, BW-BY, CA-CH, we can assign them numerically:
```{r}
cols_to_remove <- c(3:11, 14:61, 63, 75:77, 79:86)

slimmed_data <- golden_biomarkers_data$kiribati_cleandata_mergedHIES.csv %>%
  mutate(
    dob = as.Date(.data[[col_dob]], format = "%d/%m/%Y"),
    interview_date = as.Date(.data[[col_int_date]], format = "%d/%m/%Y"),
    age_in_months = dob %--% interview_date / months(1)
  ) %>%
  filter(age_in_months > 0) %>%
  select(-all_of(cols_to_remove), -dob, -interview_date, uid = hhld_id_hm_basic__id, hdl = L1_q61_hdl) %>%
  select(
    -L1_q221_agemonths, -L1_q254_confidence_in_date_of_birth,
    -dob_corrected, -hm_basic__id, L0_q1_hh_id, -L1_q65_metabolicphoto_caption,
    -time_created, -L1_q251_notes, -L0_q2_hh_id, -L1_q5_id_pic_caption,
    -interview__key, -gender_corrected, -deleted, -hhld_id_corrected, -L0_q1_hh_id,
    -agedays_corrected, -agemonths_corrected, -ageyears_corrected,
    -name_HIES, -DOB_HIES, -L1_q65_metabolicphoto, -L1_q59_diabetes, gender = gender_HIES
    ) %>%
  select(-contains("Scaled_rank"), -contains("L1_")) %>%
  mutate(uid = purrr::map_chr(uid, digest::digest)) %>%
  mutate(across(where(is.character), as.factor)) %>%
  rename_with(~ str_remove(., "_corrected"), ends_with("_corrected")) %>%
  rename(temperature = temperature_correctedF) %>%
  mutate(uid = as.character(uid))
```

Mental note: we are introducing HDL cholesterol here manually. Whereas
most of our variables come from the "`[VAR]_corrected`" columns,
HDL is calculated in one of the columns with the `L1_` prefix at 
position 62. So while we are removing all the `L1_` columns, we 
keep this one and rename it to `hdl` for clarity.

The `slimmed_data` object is a clean dataset of biomarkers and minimal demographics.
As you can see above, we removed numerous columns that were identifiable in
anyway, keeping only age group, age in months, and gender. Importantly,
we also hashed the unique identifier for each person to further de-identify
the data. However, this removes the ability to perform hierarchical analyses
due to the fact that each record is now unique (ie no repeated measures).

```{r}
skimr::skim(slimmed_data)
```

To make a small data set of age, sex and location (note that lat and long were largely missing data):

```{r}
small_data <- golden_biomarkers_data$kiribati_cleandata_mergedHIES.csv %>%
  mutate(
    dob = as.Date(.data[[col_dob]], format = "%d/%m/%Y"),
    interview_date = as.Date(.data[[col_int_date]], format = "%d/%m/%Y"),
    age_in_months = dob %--% interview_date / months(1)
  ) %>%
  filter(age_in_months > 0) %>%
  select( 
    uid = hhld_id_hm_basic__id,
    age_in_months, 
    gender = L1_q3_gender,     community = L1_q125_which_community_are_you_in
  ) %>%
  mutate(uid = purrr::map_chr(uid, digest::digest)) %>%
  mutate(across(where(is.character), as.factor)) %>%
  mutate(uid = as.character(uid))
```

```{r}
skimr::skim(small_data)
```

Now we can use the `codebook` package to create a data dictionary for these datasets.

```{r}
slim_labels <- list(
  uid = "Unique Identifier (note that repeated measures per person have been removed to de-identify data)",
  age_group = "Age Group",
  fever = "Fever recorded",
  BloodPressure_status = "Blood Pressure Status",
  Glucose_status = "Glucose Status",
  anemia_status = "Anemia Status",
  cholesterol_status = "Cholesterol Status",
  triglycerides_status = "Triglycerides Status",
  diabetes_status = "Diabetes Status",
  gender = "Gender",
  age_in_months = "Age in Months at time of survey",
  ldl = "LDL Cholesterol",
  hdl = "HDL Cholesterol",
  tc = "Total Cholesterol",
  pulse = "Pulse",
  glucose = "Glucose",
  triglycerides = "Triglycerides",
  systolic = "Systolic Blood Pressure",
  diastolic = "Diastolic Blood Pressure",
  temperature = "Temperature in Fahrenheit",
  hemoglobin = "Hemoglobin Level",
  hematocrit = "Percentage of red blood cells",
  diabetes = "HbA1c level"
)

var_label(slimmed_data) <- slim_labels

metadata(slimmed_data)$name <- "Slimmed biomarker dataset"
metadata(slimmed_data)$description <- paste(
  "De-identified slimmed biomarker dataset.",
  "Unique identifier is uid; repeated measures per person have been removed."
)
```

```{R}
codebook_table(slimmed_data) %>%
  write_csv(here("sandbox", "JZM_collab", "data_dictionary_slimmed_data.csv"))

write_csv(slimmed_data, here("sandbox", "JZM_collab", "slimmed_data.csv"))
```

```{r}
small_labels <- list(
  uid = "Unique Identifier (note that multiple measures per person have been removed to de-identify data)",
  age_in_months = "Age in Months at time of survey",
  gender = "Gender",
  community = "Community of data collection"
)

var_label(small_data) <- small_labels
metadata(small_data)$name <- "Small de-identified dataset"
metadata(small_data)$description <- paste(
  "Small de-identified dataset with uid, age, gender, and community variables.",
  "Multiple measures per person have been removed."
)
```

```{R}
codebook_table(small_data) %>%
  write_csv(here("sandbox", "JZM_collab", "data_dictionary_small_data.csv"))

write_csv(small_data, here("sandbox", "JZM_collab", "small_data.csv"))
```
