---
title: "Stage Inputs"
execute:
  eval: false
engine: markdown
---

```{r}
knitr::opts_chunk$set(eval = FALSE)
```

Before running the input-provenance pipeline, stage the source materials at the
repository input boundary. The pipeline has exactly two prerequisites:

- `stagecoach_manifest.yml` at the repository root; and
- the staged file inventory under `data/inputs/`.

The manifest records the staging context, while `data/inputs/` contains the
files that are actually available to this project.

## Targets declared in this stage

The project-root `_targets.R` tangles this notebook together with the curation
and release-packaging notebooks. Each assignment in the disabled R chunks below
becomes one target; the chunks are disabled so that reading or rendering the
Manual never inspects protected inputs. This stage declares the two file targets
that form the input boundary.

```{r}
#| format: file
pipeline_stagecoach_manifest_path <- here::here("stagecoach_manifest.yml")
```

```{r}
#| format: file
pipeline_raw_input_files_paths <- {

  # link the manifest to the inputs
  stopifnot(file.exists(pipeline_stagecoach_manifest_path))
  input_dir <- here::here("data", "inputs")
  if (!dir.exists(input_dir)) {
    stop("Staged input directory is missing: ", input_dir)
  }

  inputs <- list.files(
    input_dir,
    recursive = TRUE,
    full.names = TRUE,
    include.dirs = FALSE,
    no.. = TRUE
  )
  if (!length(inputs)) {
    stop("No staged input files are available in: ", input_dir)
  }

  inputs
}
```

:::{.manual-prereq title="Get Stagecoach"}

As always, we use `stagecoach` to create a modular method for loading
and unloading input data at will. Ensure you have `stagecoach`
installed — typically as a tool —  with `uv`.

Importantly, I don't strongly believe this has to be done in the container,
and don't have a lot of time to retroactively add it, so I'm doing
this from the regular environment, and `uv` manages the Python version
of `stagecoach`.

```sh
# uv tool install git+https://github.com/GoldenPlanetaryHealthLab/stagecoach.git

uv tool list
```

:::


:::{.manual-procedure title="Stage Data"}

`stagecoach` can walk you through the actual steps, involving
1. `hail`, which drops a manifest in the root directory,

2. `inspect`, which helps ensure your modifications are
valid and your data access is appropriate,

3. `stage`, which runs the necessary functionality to bring
your data into the workspace.

:::

For this project, out input data is part of the Google Drive Rclone tree.

:::{.manual-check title="Data Staging Complete"}

```{bash}
data_dir=data/inputs

if [ "$(ls -A $data_dir)" ]; then
  echo "$data_dir is not empty." ✅
else
  echo "$data_dir is empty" ❌
fi

[ -f stagecoach_manifest.yml ] && echo "Manifest exists" || echo "Manifest not found"
```

:::
