---
title: "Curate Release"
execute:
  eval: false
engine: markdown
---

```{r}
knitr::opts_chunk$set(eval = FALSE)
```

This page is the source of truth for the Kiribati biomarker release curation
graph. Its target declarations are tangled into the DAG; it does not source an
exported functions file. Source data remain under `data/inputs/`; only curated
artifacts are written to `data/outputs/curated/`.

## Initialize the root targets script

Before building the notebook pipeline, initialize the standard targets script
from the repository root. The following call is shown for the operator to run
there; it is deliberately non-executable in this manual.

```r
# run this outside of the pipeline to establish targets use
targets::use_targets()
```

This is official `{targets}` functionality, not an artifact produced by the
Manual's `emit` profile. Confirm that `_targets.R` is available at the
repository root before continuing. It is the project's one target script: as
the pipeline grows, edit it to add the notebooks in `notebooks/pipeline/` that
should participate in the pipeline. Do not create a second `_targets.R` under
`pipeline/`.

::: {.manual-prereq title="Confirm The Curation Inputs"}

Stage exactly one Golden et al. `kiribati_cleandata_mergedHIES.csv` export and
one `village_metadata` CSV containing `village`, `island`, `village_id`,
and `rururb`. The pipeline fails rather than arbitrarily choosing among
multiple matches.

:::

## Curation rules migrated from the legacy vignette

The release derives age in months from interview date and date of birth, drops
nonpositive derived ages, and replaces the source identifier with its MD5
digest. The small release retains only approved `village_id` and `rururb`;
it does not contain literal community or island names. Thus the approved
metadata, rather than this pipeline, supplies the non-reversible location key (i.e., the data remains non-identifiable outside of the pipeline operator).

## Simple Pipeline

```{r}
library(dplyr)
library(stringr)
library(readr)
library(readxl)
library(lubridate)
```

Our simple pipeline takes the raw data and modifies required
column names for release. No other major curation happens here.

Notably, this is separate from Manual operation —  it doesn't require
a procedure block because it is a `targets` managed stage.

First, list the biomarker columns we got from CG:

```{r}
biomarker_output_columns_vec <- c(
  "DOB_HIES", "date_of_interview", "hhld_id_hm_basic__id", "L1_q61_hdl",
  "ldl_corrected", "tc_corrected", "pulse_corrected", "glucose_corrected",
  "triglycerides_corrected", "systolic_corrected", "diastolic_corrected",
  "temperature_correctedF", "hemoglobin_corrected", "hematocrit_corrected",
  "diabetes_corrected", "age_group", "fever", "BloodPressure_status",
  "Glucose_status", "anemia_status", "cholesterol_status",
  "triglycerides_status", "diabetes_status", "gender_HIES", "L1_q3_gender",
  "L1_q125_which_community_are_you_in", "L1_q124_which_island_are_we_on"
)
```

Then, the village columns; these are used to link the participant to
their location and mask their identity.

```{r}
village_metadata_output_columns_vec <- c("village", "island", "village_id", "rururb")
```


We create a small function here to convert an identifier to a hash,
preserving PHI.

```{r}
hash_identifier <- function(identifier) {
  vapply(as.character(identifier), digest::digest, character(1))
}
```


Here we create a mapping from the health data to human-readable labels:

```{r}
health_ptx_dictionary_labels_vec <- c(
  uid = "De-identified record identifier", hdl = "HDL cholesterol",
  ldl = "LDL cholesterol", tc = "Total cholesterol", pulse = "Pulse",
  glucose = "Glucose", triglycerides = "Triglycerides",
  systolic = "Systolic blood pressure", diastolic = "Diastolic blood pressure",
  temperature = "Temperature in Fahrenheit", hemoglobin = "Hemoglobin level",
  hematocrit = "Percentage of red blood cells", diabetes = "HbA1c level",
  age_group = "Age group", fever = "Fever recorded",
  BloodPressure_status = "Blood pressure status", Glucose_status = "Glucose status",
  anemia_status = "Anemia status", cholesterol_status = "Cholesterol status",
  triglycerides_status = "Triglycerides status", diabetes_status = "Diabetes status",
  gender = "Gender", age_in_months = "Age in months at interview"
)
```

Likewise, we create a mapping of labels for the participant data:
```{r}
demog_ptx_dictionary_labels_vec <- c(
  uid = "De-identified record identifier", age_in_months = "Age in months at interview",
  gender = "Gender", village_id = "Approved non-identifying village key",
  rururb = "Village rural/urban classification"
)
```

## Identify and read the approved inputs

The input inventory from the staging notebook is narrowed to one approved export
and one approved metadata file before either is read.


```{r}
biomarker_ptx_source_csv <- tibble(path=pipeline_raw_input_files_paths) %>%
    filter(str_detect(path, "Goldenetal_KI_biomarkers-main/kiribati_cleandata_mergedHIES.csv"))

biomarker_ptx_raw_df <- read.csv(biomarker_ptx_source_csv$path) %>%
    tibble()
```

```{r}
demog_ptx_village_source_csv <- tibble(path=pipeline_raw_input_files_paths) %>%
    filter(str_detect(path, "village_metadata"))

demog_ptx_village_raw_df <- read.csv(demog_ptx_village_source_csv$path) %>%
    tibble()
```


## Derive age and record source-quality evidence

Age derivation is separate from release shaping so the audit measures the same
intermediate data used by both releases.

```{r}
biomarker_ptx_age_df <- biomarker_ptx_raw_df %>% 
    mutate(
      dob = as_date(DOB_HIES, format = "%Y-%m-%d"),
      interview_date = as.Date(L1_q16_date_of_interview, format = "%d/%m/%Y"),
      age_in_months = interval(dob, interview_date) / months(1)
)
```


## Construct the two de-identified releases

The slimmed release contains clinical measures and status fields. The small
release joins approved non-identifying location metadata and omits literal
community and island names.

```{r}
biomarker_clean_df <- biomarker_ptx_age_df %>%
  filter(!is.na(age_in_months), age_in_months > 0) %>%
  transmute(
    uid = hash_identifier(hhld_id_hm_basic__id),
    hdl = L1_q61_hdl, ldl = ldl_corrected,
    tc = tc_corrected, pulse = pulse_corrected,
    glucose = glucose_corrected,
    triglycerides = triglycerides_corrected,
    systolic = systolic_corrected, diastolic = diastolic_corrected,
    temperature = temperature_correctedF,
    hemoglobin = hemoglobin_corrected, hematocrit = hematocrit_corrected,
    diabetes = diabetes_corrected, age_group = age_group,
    fever = fever, BloodPressure_status = BloodPressure_status,
    Glucose_status = Glucose_status, anemia_status = anemia_status,
    cholesterol_status = cholesterol_status,
    triglycerides_status = triglycerides_status,
    diabetes_status = diabetes_status, gender = gender_HIES,
    age_in_months
  ) %>%
  mutate(across(where(is.character), as.factor), uid = as.character(uid))
```

```{r}
demog_clean_df <- biomarker_ptx_age_df %>%
  filter(!is.na(age_in_months), age_in_months > 0) %>%
  transmute(
    uid = hash_identifier(hhld_id_hm_basic__id), age_in_months,
    gender = L1_q3_gender,
    community = L1_q125_which_community_are_you_in,
    island = L1_q124_which_island_are_we_on
  ) %>%
  left_join(
    select(demog_ptx_village_raw_df, village, island, village_id, rururb),
    by = c("community" = "village", "island" = "island")
  ) %>%
  select(uid, age_in_months, gender, village_id, rururb) %>%
  mutate(across(where(is.character), as.factor), uid = as.character(uid))
```

## Write release artifacts and data dictionaries

All externally consumed outputs are file targets with stable filenames.

```{r}
#| format: file
#slimmed_data_csv <- write_csv(
#  slimmed_data, file.path("data", "outputs", "curated", "slimmed_data.csv")
#)
```

```{r}
#| format: file
#small_data_csv <- write_csv(
#  small_data, file.path("data", "outputs", "curated", "small_data.csv")
#)
```

```{r}
#| format: file
#slimmed_data_dictionary_csv <- write_dictionary(
#  slimmed_data, slimmed_dictionary_labels,
#  file.path("data", "outputs", "curated", "data_dictionary_slimmed_data.csv")
#)
```

```{r}
#| format: file
#small_data_dictionary_csv <- write_dictionary(
#  small_data, small_dictionary_labels,
#  file.path("data", "outputs", "curated", "data_dictionary_small_data.csv")
#)
```

```{r}
#| format: file
#curation_audit_csv <- write_csv(
#  curation_audit_table,
#  file.path("data", "outputs", "curated", "curation_audit.csv")
#)
```

Render with the `emit` profile to materialize the compact orchestrator. Render
with the `pipeline` profile only after reviewing the staged inputs and rules.
