---
title: "Package Release"
execute:
  eval: false
engine: markdown
---

```{r}
knitr::opts_chunk$set(eval = FALSE)
```

This page is the source of truth for release serialization. It turns the two
curated participant-grain targets into a single Frictionless Data Package. The
package descriptor carries the data dictionaries as field descriptions, so the
release does not require separately maintained dictionary CSV files.

## Describe the biomarker resource

`create_schema()` infers field types and factor enumerations. This target adds
the release-specific descriptions and fails if the description set and curated
columns diverge.

```{r}
biomarker_ptx_dictionary_field_described_list <- {
  biomarker_description_chr <- c(
    uid = "De-identified record identifier",
    hdl = "HDL cholesterol",
    ldl = "LDL cholesterol",
    tc = "Total cholesterol",
    pulse = "Pulse",
    glucose = "Glucose",
    triglycerides = "Triglycerides",
    systolic = "Systolic blood pressure",
    diastolic = "Diastolic blood pressure",
    temperature = "Temperature in Fahrenheit",
    hemoglobin = "Hemoglobin level",
    hematocrit = "Percentage of red blood cells",
    diabetes = "HbA1c level",
    age_group = "Age group",
    fever = "Fever recorded",
    BloodPressure_status = "Blood pressure status",
    Glucose_status = "Glucose status",
    anemia_status = "Anemia status",
    cholesterol_status = "Cholesterol status",
    triglycerides_status = "Triglycerides status",
    diabetes_status = "Diabetes status",
    gender = "Gender",
    age_in_months = "Age in months at interview"
  )
  biomarker_schema_list <- frictionless::create_schema(
    biomarker_clean_df
  )
  biomarker_field_name_chr <- vapply(
    biomarker_schema_list$fields,
    `[[`,
    character(1),
    "name"
  )
  if (!setequal(names(biomarker_description_chr), biomarker_field_name_chr)) {
    stop("Biomarker field descriptions do not match the curated columns.")
  }
  biomarker_schema_list$fields <- lapply(
    biomarker_schema_list$fields,
    function(field_list) {
      field_list$description <- unname(
        biomarker_description_chr[[field_list$name]]
      )
      field_list
    }
  )

  biomarker_schema_list
}
```

## Describe the demographic resource

```{r}
demog_ptx_dictionary_field_described_list <- {
  demog_description_chr <- c(
    uid = "De-identified record identifier",
    age_in_months = "Age in months at interview",
    gender = "Gender",
    village_id = "Approved non-identifying village key",
    rururb = "Village rural/urban classification"
  )
  demog_schema_list <- frictionless::create_schema(
    demog_clean_df
  )
  demog_field_name_chr <- vapply(
    demog_schema_list$fields,
    `[[`,
    character(1),
    "name"
  )
  if (!setequal(names(demog_description_chr), demog_field_name_chr)) {
    stop("Demographic field descriptions do not match the curated columns.")
  }
  demog_schema_list$fields <- lapply(
    demog_schema_list$fields,
    function(field_list) {
      field_list$description <- unname(demog_description_chr[[field_list$name]])
      field_list
    }
  )

  demog_schema_list
}
```

## Assemble the data package

The resource names are also the stable CSV basenames written by
`write_package()`.

```{r}
release_kiribati_data_resource_packaged_list <-
  frictionless::create_package(
    list(
      name = "kiribati-biomarkers",
      title = "Kiribati Biomarkers",
      description = paste(
        "De-identified participant biomarker and demographic resources",
        "for the Kiribati Biomarkers v0.1.0 release."
      ),
      version = "0.1.0",
      resources = list()
    )
  ) |>
  frictionless::add_resource(
    resource_name = "biomarker_ptx",
    data = biomarker_clean_df,
    schema = biomarker_ptx_dictionary_field_described_list,
    title = "Participant biomarker data"
  ) |>
  frictionless::add_resource(
    resource_name = "demog_ptx",
    data = demog_clean_df,
    schema = demog_ptx_dictionary_field_described_list,
    title = "De-identified participant demographic data"
  )
```

## Publish the immutable release

Versioned releases are immutable. This target refuses to write into a nonempty
version directory, which prevents `write_package()` from retaining stale CSVs.
Changed release content requires a new version directory.

```{r}
#| format: file
release_kiribati_data_file_published_chr <- {
  release_output_directory_chr <- here::here(
    "data",
    "outputs",
    "releases",
    "v0.1.0"
  )
  release_expected_file_chr <- file.path(
    release_output_directory_chr,
    c("datapackage.json", "biomarker_ptx.csv", "demog_ptx.csv")
  )

  if (
    dir.exists(release_output_directory_chr) &&
      length(
        list.files(
          release_output_directory_chr,
          all.files = TRUE,
          no.. = TRUE
        )
      )
  ) {
    stop(
      "Release directory is nonempty and versioned releases are immutable: ",
      release_output_directory_chr
    )
  }

  dir.create(
    release_output_directory_chr,
    recursive = TRUE,
    showWarnings = FALSE
  )
  frictionless::write_package(
    release_kiribati_data_resource_packaged_list,
    directory = release_output_directory_chr
  )

  release_missing_file_chr <- release_expected_file_chr[
    !file.exists(release_expected_file_chr)
  ]
  if (length(release_missing_file_chr)) {
    stop(
      "Frictionless did not write the expected release files: ",
      paste(release_missing_file_chr, collapse = ", ")
    )
  }

  release_expected_file_chr
}
```
